### Introduction
Peer Assignment1 : "PA1_template.Rmd"

### A. Loading and preprocessing the data
```{r setoptions, echo=TRUE}
## read in the data and explore
data1 <- read.csv("activity.csv")
str(data1)
summary(data1)
class(data1)
## transform the data in date column into readable R format
convertDate <- as.Date(data1$date)
## create new data frame with convertDate
data2 <- data.frame(data1$steps, convertDate, data1$interval)
## rename colummns
colnames(data2) <- c("steps", "date", "interval")
```


### B. Mean total number of steps per day
```{r}
## create table that averages number of step per day
tableStep <- tapply(data2$step,data2$date, sum)
```

Histogram of total number of steps/day
```{r, results }
hist(tableStep, col = "blue", main="Total number of steps/day", xlab = "Number of steps")
mean(tableStep, na.rm = TRUE)
median(tableStep, na.rm = TRUE)
```
Mean number of steps: 10766
Median number of steps: 10765


### C. Average daily pattern
Create a time series of plot of 5-minute interval and average number of steps 
```{r}
plot(data2$interval, data2$steps, type= "l", xlab = "time in hours", 
    ylab = "Average number of steps/day", main = "Average number of steps across all intervals")
```

Calculate the 5-minute inteval that has the maximum number of steps

```{r}
meanStepsInterval <- aggregate(steps~interval, data = data2, mean)
head(meanStepsInterval)
## get the row number with the maximum number of steps
which.max(meanStepsInterval$steps)
## returns that row 
meanStepsInterval[104,]
```
The 5-minute interval with maximum number of steps: 835


### D. Imputting missing values
Calculate rows with missing values
```{r}
good <- complete.cases(data2)
head(good)
table(good)
```
Number of rows with missing values: 2304

### Replacing rows with missing values (NA) with the average of th t ime intverval

```{r}
library(plyr)
impute <- function(x, fun) {
  missing <- is.na(x)
  replace(x, missing, fun(x[!missing]))
}
replaceNA <- ddply(data2, ~ interval, transform, steps = impute(steps, mean))
head(replaceNA)
```

Histogram of new dataset with replaced missing values
```{r}
tableStepNA <- tapply(replaceNA$step,replaceNA$date, sum)
head(tableStepNA)
hist(tableStepNA, col = "red", main="Total number of steps/day", xlab = "Number of steps")
mean(tableStepNA)
median(tableStepNA)
```
Mean total number of steps: 10766 
Median total number of steps: 10766 

The values for mean is the same but the values for median differed very slightly between the dataset with or without missing data. 

Imputing missing data has no effect on the mean but has a tiny effect on the median for the estimates of the total daily number of steps.


### E. Differences between activity patterns in weekdays and weekends
```{r}
## convert the date factor to weekdays
convertToDate <- replaceNA$date
days <- weekdays(convertToDate)
## convert the days to return True if weekend
weekendCheck <- (days == "Saturday") | (days == "Sunday")
## create new data frame including add weekendCheck
data3 <- cbind(replaceNA, weekendCheck)
## plot with lattice
library(lattice)
## convert weekendCheck to a factor variable, and rename them to "Weekend" and "Weekday"
data3$weekendCheck <- as.factor(data3$weekendCheck)
levels(data3$weekendCheck) <- c("weekend", "weekday")
xyplot(steps~interval | weekendCheck, data = data3, type = "l", layout = c(1,2), xlab = "interval", ylab = "Number of steps")
```

